\section{The Construction}

We will now construct counterexamples to the question raised in the beginning, that is circle bundles $E \to M$ over enlargeable $M$, where the total space $E$ admits a metric of positive scalar curvature. 
In the following we will refer to orientable manifolds with non-spin universal cover as \emph{totally non-spin}.
% In order to show the existence of such a metric we will employ the following Theorem.

% \begin{theorem}[Gromov, Hanke \cite{gromov2024torsion}]
%     \label{thm:gromov_hanke}
%     Let $V$ be a closed connected oriented smooth $n$-manifold with non-spin universal cover, $n \geq 5$. Let $\phi \colon V \to B\pi_1(V)$ be the classifying map and assume that there exists a closed oriented $n$-manifold $V^\prime$, which admits a positive scalar curvature metric, as well as a map $f \colon V^\prime \to B\pi_1(V)$, such that 
%     \[
%     \phi_\ast([V]) = f_\ast([V^\prime]) \in H_n(\pi_1(V);\Z).
%     \]
%     (This includes the case $\phi_\ast([V]) = 0$.) Then $V$ admits a positive scalar curvature metric.
% \end{theorem}
For $n \geq 2$ let $N^{2n}$ be any $2n$-dimensional totally non-spin manifold. We could for example take $N:= \CP^2 \times T^{2n-4}$, which is totally non-spin by Lemma \ref{lem:totally_non_spin_lemma} and Lemma \ref{lem:cp2_not_spin} (see Appendix). Now consider the manifold
\[
M := \CP^n \# T^{2n} \# N.
\]
Since the collapse map $M \to T^{2n}$ has degree $1$ and $T^{2n}$ is enlargeable, it follows from Lemma \ref{lem:constructing_enlargeable_mnf} that $M$ is also enlargeable. We will now construct a circle bundle $E \to M$ as follows. Let 
\[
c_1 \in H^2(M;\Z) \cong H^2(\CP^n;\Z) \oplus H^2(T^{2n};\Z) \oplus H^2(N;\Z)
\]
be the cohomology class corresponding to the first chern class $\alpha \in H^2(\CP^n;\Z)$ of the canonical bundle $S^{2n+1} \to \CP^n$. We then define $E \to M$ as the unique $S^1$-principal bundle with first chern class equal to $c_1$. Now this is quite an abstract definition, so let us develop an intuition for how this bundle looks like. We claim that $E$ is trivial over the summand $T^{2n} \# N$ and is given by the restriction of the canonical bundle $S^{2n+1} \to \CP^n$ over the $\CP^n$ summand. To make this precise let 
\[
i_1 \colon (T^{2n} \# N) - D^{2n} \to M
\]
be the inclusion. Then
\[\begin{tikzcd}
	{H^2(M; \Z)} & {H^2((T^{2n} \# N) - D^{2n};\Z) } \\
	{H^2(\CP^n;\Z) \oplus H^2(T^{2n} \# N;\Z)} & {H^2(T^{2n} \# N;\Z) }
	\arrow["{i_1^\ast}", from=1-1, to=1-2]
	\arrow["\cong", from=2-1, to=1-1]
	\arrow["{\mathrm{proj.}}", from=2-1, to=2-2]
	\arrow["\cong", from=2-2, to=1-2]
\end{tikzcd}\]
commutes and therefore $i_1^\ast (c_1) = 0$. Again using the fact that $S^1$-principal bundles are uniquely determined by their first chern class we see that $i_1^\ast E$ is indeed trivial. Similiarly if we denote by 
\[
i_2 \colon \CP^n - D^{2n} \to M, \quad j \colon \CP^{n}-D^{2n} \to \CP^{n}
\]
the inclusions, the diagram
\[\begin{tikzcd}
	{H^2(M ;\Z)} & {H^2(\CP^n - D^{2n};\Z) } \\
	{H^2(\CP^n;\Z) \oplus H^2(T^{2n} \# N;\Z)} & {H^2(\CP^n;\Z) }
	\arrow["{i_2^\ast}", from=1-1, to=1-2]
	\arrow["\cong", from=2-1, to=1-1]
	\arrow["{\mathrm{proj.}}", from=2-1, to=2-2]
	\arrow["\cong", "j^\ast"', from=2-2, to=1-2]
\end{tikzcd}\]
commutes and thus $i_2^\ast(c_1) = j^\ast(\alpha)$. So the first chern classes of $i_2^\ast E$ and $j^\ast S^{2n+1}$ agree, hence they must be isomorphic bundles. 

With this intuition in mind we will now show the following two claims.

\begin{claim}
	\label{claim1}
$E$ is totally non-spin.
\end{claim}

\begin{claim}
	\label{claim2}
The projection $p \colon E \to M$ induces an isomorphism on $\pi_1$.
\end{claim}

Granting these claims for now, let us see how to conclude the existence of a positive scalar curvature metric on $E$. The disk bundle $DE \to M$ corresponding to $E$ defines an oriented null-cobordism of $E$. Moreover, since $DE \to M$ is a homotopy equivalence it follows from Claim \ref{claim1} that the inclusion $E \to DE$ is an isomorphism on $\pi_1$. So together with Claim \ref{claim2} the following Theorem applies.

\begin{theorem}
	Let $X^n$ be a closed connected totally non-spin manifold of dimension $n \geq 5$. Suppose there exists an oriented null-cobordism $W$ of $X$, such that the inclusion $X \to W$ induces an isomorphism on $\pi_1$. Then $X$ admits a metric of positive scalar curvature.
\end{theorem}

We will give a proof of this theorem further below, which is essentially an application of the well known Gromov-Lawson surgery principle (SOURCE). Let us now proof Claim \ref{claim1} and Claim \ref{claim2}.

% Granting these Claims for now let us see why they suffice for the assumptions of Theorem \ref{thm:gromov_hanke} to be satisfied.  
% For this we would like to show that the classifying map $\phi \colon E \to B\pi_1(E)$ maps $[E]$ to $0 \in H_{2n+1}(\pi_1(E);\Z)$, which can be seen by using Claim \ref{claim2} as follows. Let $q \colon DE \to M$ be the disk bundle corresponding to $E$. Moreover let $\Phi \colon DE \to B\pi_1(DE)$ be the classifiying map for $DE$. Then the diagram 
% \[\begin{tikzcd}
% 	DE & {B\pi_1(DE)} \\
% 	E & {B\pi_1(E)}
% 	\arrow["\Phi", from=1-1, to=1-2]
% 	\arrow["\iota", from=2-1, to=1-1]
% 	\arrow["\phi", from=2-1, to=2-2]
% 	\arrow["{B\pi_1(\iota)}"', from=2-2, to=1-2]
% \end{tikzcd}\]
% commutes (hier noch BegrĂĽndung rein), where the vertical maps are (induced by) the inclusion $\iota \colon E \to DE$. We claim that $B\pi_1(\iota)$ is an isomorphism. This follows from commutativity of  
% \[\begin{tikzcd}[column sep = small]
% 	{B\pi_1(E)} && {B\pi_1(DE)} \\
% 	& {B\pi_1(M)}
% 	\arrow["{B\pi_1(\iota)}", from=1-1, to=1-3]
% 	\arrow["{B\pi_1(p)}"',"\cong", from=1-1, to=2-2]
% 	\arrow["{B\pi_1(q)}","\cong"', from=1-3, to=2-2]
% \end{tikzcd}\]
% where $B\pi_1(q)$ is an isomorphism by homotopy invariance and $B\pi_1(p)$ is an isomorphism by Claim \ref{claim2}. Thus $\phi$ factors over $\iota$. Since $DE$ is an oriented null-cobordism of $E$ it follows that $\iota_\ast[E] = 0 \in H_{2n+1}(DE;\Z)$ and so in particular $\phi_\ast [E] = 0\in H_{2n+1}(\pi_1(E);\Z)$ as we wanted to show. So together with Claim \ref{claim1} all assumptions of Theorem \ref{thm:gromov_hanke} are satisfied, showing that $E$ admits a metric of positive scalar curvature. 

\begin{proof}[Proof of Claim \ref{claim1}]
	Let $i \colon N- D^{2n} \to M$ be the inclusion. We have already established above, that $E$ is trivial over the $T^{2n} \# N$ summand and so in particular
	\[
	i^\ast E \cong (N- D^{2n}) \times S^1.
	\]
	From Lemma \ref{lem:totally_non_spin_lemma} we conclude that $N_0 := N - D^{2n}$ and consequently also $(N- D^{2n}) \times S^1$ are totally non-spin. 
	% Then the universal cover of $i^\ast E$ is given by $\tilde{N}_0 \times \R$, where $\tilde{N}_0$ is the universal cover of $N_0$. Let $\pi_{\tilde{N}_0} \colon \tilde{N}_0 \times \R \to \tilde{N}_0$ and $\pi_\R \colon \tilde{N}_0 \times \R \to \R$ be the projections. We compute 
	% \[
	% w_2(T(\tilde{N}_0 \times \R)) = w_2(\pi_{\tilde{N}_0}^\ast T\tilde{N}_0 \oplus \pi_\R^\ast T\R) = \pi_{\tilde{N}_0}^\ast w_2(T\tilde{N}_0).
	% \]
	% Since $\pi_{\tilde{N}_0}$ is a homotopy equivalence and $\tilde{N}_0$ is non-spin it follows that $w_2(T(\tilde{N}_0 \times \R)) \neq 0$, such that $i^\ast E$ is totally non spin. 
	Therefore $E$ contains a totally non-spin submanifold of codimension $0$ and so again Lemma \ref{lem:totally_non_spin_lemma} implies that $E$ is totally non-spin.
\end{proof}

\begin{proof}[Proof of Claim \ref{claim2}]
	Let $j \colon \CP^{n} - D^{2n}$ be the inclusion. Consider the following commutative diagram, where the rows are taken from the long exact sequences of homotopy groups of the respective fibrations and the vertical maps are induced by inclusions: 
	\[\begin{tikzcd}
	{\pi_2(M)} & {\pi_1(S^1)} & {\pi_1(E)} \\
	{\pi_2(\CP^n - D^{2n})} & {\pi_1(S^1)} & {\pi_1(j^\ast E)} \\
	{\pi_2(\CP^n)} & {\pi_1(S^1)} & {\pi_1(S^{2n+1})}
	\arrow["{\partial_1}", from=1-1, to=1-2]
	\arrow[from=1-2, to=1-3]
	\arrow[from=2-1, to=1-1]
	\arrow["{\partial_2}", from=2-1, to=2-2]
	\arrow["\cong"', from=2-1, to=3-1]
	\arrow["{\mathrm{id}}", from=2-2, to=1-2]
	\arrow[from=2-2, to=2-3]
	\arrow["{\mathrm{id}}"', from=2-2, to=3-2]
	\arrow[from=2-3, to=1-3]
	\arrow[from=2-3, to=3-3]
	\arrow["{\partial_3}", from=3-1, to=3-2]
	\arrow[from=3-2, to=3-3]
\end{tikzcd}\]
The vertical map on the lower left is an isomorphism, because both $\CP^n$ and $\CP^n - D^{2n}$ are simply connected and so by the Hurewicz Theorem
\[\begin{tikzcd}
	{\pi_2(\CP^n-D^{2n})} & {\pi_2(\CP^n)} \\
	{H_2(\CP^n-D^{2n};\Z)} & {H_2(\CP^n;\Z)}
	\arrow[from=1-1, to=1-2]
	\arrow["\cong", from=1-1, to=2-1]
	\arrow["\cong", from=1-2, to=2-2]
	\arrow["\cong", from=2-1, to=2-2]
\end{tikzcd}\]
commutes and the vertical maps are isomorphisms. We now claim that $\partial_1$ is surjective. For this it suffices to show surjectivity of $\partial_2$, which in turn is equivalent to surjectivity of $\partial_3$. But this is clear by exactness of the rows and $\pi_1(S^{2n+1}) = 0$. Consequently
\[
\pi_1(S^1) \to \pi_1(E)
\]
is the zero map and therefore we conclude from exactness of
\[
\pi_1(S^1) \to \pi_1(E) \to \pi_1(M) \to \pi_0(S^1),
\]
that $\pi_1(E) \to \pi_1(M)$ is indeed an isomorphism.
\end{proof}

The above construction now gives examples for circle bundles with positive scalar curvature over enlargeable manifolds in all even dimensions $\geq 4$. To also get examples in odd dimensions let $E \to M$ be as above and let $\pi \colon M \times S^1 \to M$ be the projection. Then the pullback bundle $\pi^\ast E \cong E \times S^1$ clearly admits a metric of positive scalar curvature and by Lemma \ref{lem:constructing_enlargeable_mnf} the base space $M \times S^1$ is still enlargeable.