\appendix
\section{Appendix}

The following lemma collects a few basic facts about totally non-spin manifolds.

\begin{lemma}
    \label{lem:totally_non_spin_lemma} 
    \begin{enumerate}[label=(\roman*)]
        \item[] 
        \item The product of a totally non-spin manifold with any non-empty manifold is totally non-spin. \label{eq:totally_non_spin_lemma1}
        \item A manifold containing a totally non-spin submanifold of codimension $0$ is totally non-spin. \label{eq:totally_non_spin_lemma2}
        \item If $M$ is totally non-spin and of dimension $n \geq 4$, then $M - D^n$ is totally non-spin for any embedded disk $D^n \subseteq \mathrm{int}(M)$. \label{eq:totally_non_spin_lemma3}
    \end{enumerate}
\end{lemma}

\begin{remark}
    All statements of the lemma remain valid if ``totally non-spin'' is replaced by ``non-spin''. The corresponding proofs are just simpler versions of the arguments given below.
\end{remark}

\begin{proof}%[Proof of Lemma \ref{lem:totally_non_spin_lemma}]
    For \ref{eq:totally_non_spin_lemma1} let $M_1$ and $M_2$ be manifolds with $M_1$ totally non-spin. Then the universal cover of the product $M_1 \times M_2$ is given by the product $\tilde{M_1} \times \tilde{M_2}$ of the respective universal covers. Denote by 
    \[
    \pi_1 \colon \tilde{M_1} \times \tilde{M_2} \to \tilde{M_1}, \; \;  \pi_2 \colon \tilde{M_1} \times \tilde{M_2} \to \tilde{M_2}
    \]
    the projections and let $\iota \colon \tilde{M_1} \to \tilde{M_1} \times \tilde{M_2}$ be an inclusion for some choice of basepoint in $\tilde{M_2}$. We compute 
    \begin{align*}
    \iota^\ast w_2(T(\tilde{M_1} \times \tilde{M_2})) & = \iota^\ast w_2(\pi_1^\ast T \tilde{M_1} \oplus \pi_2^\ast T \tilde{M_2}) \\
    & = w_2((\pi_1 \circ \iota)^\ast T\tilde{M_1} + (\pi_2 \circ \iota)^\ast T\tilde{M_2} ) \\
    & = w_2(T \tilde{M_1}) \neq 0,
    \end{align*}
    where we used that $\pi_2 \circ \iota$ is constant, such that $(\pi_2 \circ \iota)^\ast T\tilde{M_2}$ is trivial. Thus $w_2(T(\tilde{M_1} \times \tilde{M_2})) \neq 0$.

    For \ref{eq:totally_non_spin_lemma2} suppose that $N \subseteq M$ is some totally non-spin codimension $0$ submanifold of the manifold $M$. Then the inclusion $\iota \colon N \to M$ lifts to the universal covers:
    \[\begin{tikzcd}
	{\tilde{N}} & {\tilde{M}} \\
	N & M
	\arrow["{\tilde{\iota}}", from=1-1, to=1-2]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-2, to=2-2]
	\arrow["\iota", from=2-1, to=2-2]
    \end{tikzcd}\]
    Since $\tilde{\iota}$ is a local diffeomorphism, the differential $d \tilde{\iota} \colon T\tilde{N} \to T\tilde{M}$ is a fiber-wise isomorphism, such that the following is a pullback diagram
    \[\begin{tikzcd}
	{T\tilde{N}} & {T\tilde{M}} \\
	{\tilde{N}} & {\tilde{M}}
	\arrow["{d \tilde{\iota}}", from=1-1, to=1-2]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-2, to=2-2]
	\arrow["{\tilde{\iota}}", from=2-1, to=2-2]
    \end{tikzcd}\]
    Therefore $\tilde{\iota}^\ast w_2(T\tilde{M}) = w_2(T\tilde{N}) \neq 0$ and so $w_2(T\tilde{M}) \neq 0$.

    For \ref{eq:totally_non_spin_lemma3} assume without of loss generality that $M$ is connected. Denote by $B^n := \mathrm{int}(D^n)$ the open ball. Then inclusion 
    \[
    M - D^n \to M - B^n
    \]
    is a homotopy equivalence from which we conclude (using analogous arguments as for \ref{eq:totally_non_spin_lemma2}) that $M - D^n$ is totally non-spin if and only if $M- B^n$ is totally non-spin. We will now show this for the latter. Let $p \colon \tilde{M} \to M$ be the universal cover of $M$ and write $M_0 := M - B^n$. We claim that the universal cover of $M_0$ agrees with the covering 
    \[
    \hat{M}_0 = \tilde{M} - p^{-1}(B^n) \xrightarrow{p} M_0.
    \]
    First of all $\hat{M}_0$ is connected as $n \geq 2$. We now have to show that $\hat{M}_0$ is simply connected. Let $F = p^{-1}(x)$ be some fiber of $p$. From the classification of coverings we see that 
    \[
    p^{-1}(B^n) \cong B^n \times F
    \]
    since $B^n$ is simply connected. If the fiber $f$ is finite we can apply the Seifer-van-Kampen Theorem to conclude that $\tilde{M}$ with a single disk $B^n$ removed is still simply connected, such that the statement follows inductively. If the fiber is not finite we can argue more generally as follows. Let $\phi \colon S^1 \to \hat{M}_0$ be some smooth loop. Certainly we can contract this loop in $\tilde{M}$, that is we can find some map $\Phi \colon D^2 \to \tilde{M}$ restricting to $\alpha$ on the boundary. We now show that $\Phi$ is homotopic rel$\,S^1$ to some map $\Phi_1 \colon D^2 \to \tilde{M}$, such that the image of $\Phi_1$ is already contained in $\tilde{M}_0$, from which the claim follows. To construct such a homotopy let $0 \in B^n \subseteq \tilde{M}$ be the center of the embedded disk. After a preliminary homotopy we may assume that $\Phi$ is transversal to $p^{-1}(0)$. Since $n \geq 3$ this is equivalent to requiring that $\Phi$ does not intersect $p^{-1}(0)$. Now note that $S^{n-1}$ is a strong deformation retract of $D^n - 0$ and applying such a deformation retraction on each disk $D^n - 0 \subseteq p^{-1}(D^n - 0)$ shows that $\hat{M}_0 = \tilde{M} - p^{-1}(B^n)$ is a strong deformation retract of $\tilde{M} - p^{-1}(0)$. Let 
    \[
    r \colon \tilde{M} - p^{-1}(0) \to \hat{M}_0
    \]
    be such a strong deformation retraction and let 
    \[
    r_t \colon \tilde{M} - p^{-1}(0) \to \tilde{M} - p^{-1}(0)
    \]
    be a homotopy from the identity to $r$ (composed with the corresponding inclusion). Then $\Phi_t := r_t \circ \Phi$ defines a homotopy rel$\,S^1$ from $\Phi$ to some map $\Phi_1$ whose image is contained in $\hat{M}_0$. 
    % Since $\Phi_1^\prime(S^1) = \phi(S^1)$ is contained in the interior $\hat{M}_0$ of this manifold we can take a further homotopy rel$\,S^1$ from $\Phi_1^\prime$ to a map $\Phi_1$ whose image is contained in $\hat{M}_0$. 
    This shows the claim and hence the universal cover of $M_0$ is given by 
    \[
    \tilde{M}_0 = \hat{M}_0 = \tilde{M} - p^{-1}(B^n) \cong \tilde{M} - (B^n \times F)
    \]
    Finally the Mayer-Vietoris sequence of the pushout 
    \[\begin{tikzcd}
	{S^{n-1} \times F} & {D^n \times F} \\
	{\tilde{M} - (B^n \times F)} & {\tilde{M}}
	\arrow[from=1-1, to=1-2]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-2, to=2-2]
	\arrow["\iota", from=2-1, to=2-2]
    \end{tikzcd}\]
    where all maps are inclusions, then shows that $\iota$ is an isomorphism on $H^2(-;\Z_2)$. Here we need $n \geq 4$. Therefore
    \[
    w_2(T\tilde{M}_0) = \iota^\ast w_2(T\tilde{M}) \neq 0.
    \]
\end{proof}

In the construction in section 2 we also made use of the following fact.

\begin{lemma} \label{lem:cp2_not_spin}
    $\C P^2$ is totally non-spin.
\end{lemma}